---
title: "Effect on Health & Economic due to Storm"
author: "prabhu manickavelu"
date: "24 November 2016"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Storms effect the health of the population and economy of that region.

Hear we'll explore the NOAA Storm Database to which type of event effect the health of population the most as well as the economy. Across the United States.

```{r cache=TRUE, load_data}
if(!file.exists("StormData.csv.bz")){
    print("not there")
    download.file("https://d396qusza40orc.cloudfront.net/repdata%2Fdata%2FStormData.csv.bz2","StormData.csv.bz")   
}
StormData <- read.csv("StormData.csv.bz")

```
## Data Processing

### FATALITIES & INJURIES
Consolidating injuries and fatalities then getting the sum of it with respect to event type to find the event which most injuries and fatalities.
```{r}
    StormData$FI <- StormData$FATALITIES+StormData$INJURIES
    injuries <- tapply(StormData$FI, StormData$EVTYPE, sum)
```

### Property and Crop Damage
Simple function to convert the amount
```{r}
    conver_text <- function(vals, mags){
        # keeping ""  "-" "?" "+" "0" as 0 presision
        # converting billions
        vals[mags %in% c("B","b")] <- vals[mags %in% c("B","b")] * 1000000000
        # converting milions
        vals[mags %in% c("M","m")] <- vals[mags %in% c("M","m")] * 1000000
        # converting thousands
        vals[mags %in% c("K","k")] <- vals[mags %in% c("K","k")] * 1000
        # converting h,H to hundreds
        vals[mags %in% c("H","h")] <- vals[mags %in% c("H","h")] * 100
        # converting the presision values
        vals[mags %in% c("1","2","3","4","5","6","7","8")] <- vals[mags %in% c("1","2","3","4","5","6","7","8")] * 10^as.numeric(mags[mags %in% c("1","2","3","4","5","6","7","8")])
        return(vals)
    }
```
calculating the total damage by adding property and crop damage
```{r}
    StormData$dmg <- conver_text(StormData$CROPDMG, StormData$CROPDMGEXP) + conver_text(StormData$PROPDMG, StormData$PROPDMGEXP)
    damage <- tapply(StormData$dmg, StormData$EVTYPE, sum)
```


## Results
### FATALITIES & INJURIES
The max amount of popluation who were inhured or had fatalities were in `r names(which.max(injuries))`

The Below graph represents the findings 
```{r message=FALSE, warning=FALSE}
    library(plotly)
    
    plot_ly(data.frame(fi = injuries[injuries>0], evntyp=names(injuries[injuries>0])), labels = ~evntyp, values = ~fi, type = 'pie') %>% layout(title = 'Injuries and Fatalities due to storms Events', xaxis = list(showgrid = FALSE, zeroline = FALSE, showticklabels = FALSE), yaxis = list(showgrid = FALSE, zeroline = FALSE, showticklabels = FALSE))

```

### Damage
The max economic damage was through `r names(which.max(damage))`

The Below graph represents the findings 
```{r message=FALSE, warning=FALSE}
    plot_ly(data.frame(fi = damage[damage>0]/1000000, evntyp=names(damage[damage>0])), labels = ~evntyp, values = ~fi, type = 'pie') %>% layout(title = 'Economic Damage due to storms Events', xaxis = list(showgrid = FALSE, zeroline = FALSE, showticklabels = FALSE), yaxis = list(showgrid = FALSE, zeroline = FALSE, showticklabels = FALSE))
```


